---
title: "More Loops Practice"
author: "Kevin Patyk"
date: '2022-09-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1: Print First 10 natural numbers using while loop

```{r}
i = 1
while( i <= 10){
  print(i)
  i = i + 1
}
```

## Exercise 2: Print the following pattern

* `1`
* `1 2`
* `1 2 3`
* `1 2 3 4`
* `1 2 3 4 5`

```{r}
stor = c()
for(i in 1:5){
  for(j in 1:i){
    stor = c(stor, j)
  }
  print(stor)
  stor = c()
}
```

## Exercise 3: Calculate the sum of all numbers from 1 to a given number

```{r}
N = 10
sum = 0
for(i in 1:N){
  sum = sum + i
}
print(sum)
```

## Exercise 4: Write a program to print multiplication table of a given number

```{r}
num = 3
stor = 0
for(i in 1:10){
  stor = stor + num
  print(stor)
}
```

Or:

```{r}
num = 2
for(i in 1:10){
  product = num * i
  print(product)
}
```

## Exercise 5: Display numbers from a list using loop

Write a program to display only those numbers from a list that satisfy the following conditions:

* The number must be divisible by five
* If the number is greater than 150, then skip it and move to the next number
* If the number is greater than 500, then stop the loop

Given:

`numbers = [12, 75, 150, 180, 145, 525, 50]`

Expected output:

`75`
`150`
`145`

```{r}
numbers = c(12,75,150,180,145,525,50)
for(i in numbers){
  if(i > 500){
    break()
  } else if(i > 150 | i %% 5 != 0){
    next()
  } else{
    print(i)
  }
}
```

## Exercise 6:

Write a program to count the total number of digits in a number using a while loop.

For example, the number is 75869, so the output should be 5.

```{r}
number = 75869
counter = 0
while(number != 0){
  number = floor(number / 10)
  counter = counter + 1
}

print(counter)
```

## Exercise 7: Print the following pattern

Write a program to use for loop to print the following reverse number pattern:

* `1 2 3 4 5`
* `1 2 3 4`
* `1 2 3`
* `1 2`
* `1`

```{r}
stor = c()
k = 6
for(i in 1:5){
  for(j in 5:(k-i)){
    stor = c(stor, j)
  }
  print(stor)
  stor = c()
}
```

## Exercise 8: Print list in reverse order using a loop

* Given:

`list1 = [10, 20, 30, 40, 50]`

```{r}
list1 = c(10, 20, 30, 40, 50)
newlist = c()
for(i in length(list1):1){
  newlist = c(newlist, list1[i])
}

print(newlist)
```

## Exercise 9: Display numbers from -10 to -1 using for loop

```{r}
stor <- c()
for(i in -10:-1){
  stor <- c(stor, i)
}
print(stor)
```

## Exercise 10: Use else block to display a message “Done” after successful execution of for loop

```{r}
n = 5
for(i in 1:n){
  print(i)
} 
if(i == n){
  print("Done")
}
```

## Exercise 11: Write a program to display all prime numbers within a range

**Note**: A Prime Number is a number that cannot be made by multiplying other whole numbers. A prime number is a natural number greater than 1 that is not a product of two smaller natural numbers

Examples:
* 6 is not a prime mumber because it can be made by 2×3 = 6
* 37 is a prime number because no other whole numbers multiply together to make it.

Given:

* `start = 25`
* `end = 50`

```{r}
n = 50
    for (j in 2:n) {
      f = 1
      i = 2
      n = j
      while (i <= n / 2) {
        if (n %% i == 0) {
          f = 0
          break
        }
        i = i + 1
      }
      if (f == 1) {
        print(paste("Number is prime :", n))
      }
    }
```

## Exercise 12: Display Fibonacci series up to 10 terms

The Fibonacci Sequence is a series of numbers. The next number is found by adding up the two numbers before it. The first two numbers are 0 and 1.

For example, 0, 1, 1, 2, 3, 5, 8, 13, 21. The next number in this series above is 13+21 = 34.

Expected output: 

* `0  1  1  2  3  5  8  13  21  34`

```{r}
N = 10
n1 = 0
n2 = 1
for(i in 1:N){
  print(n1)
  result = n1 + n2
  n1 = n2
  n2 = result
}
```

## Exercise 13: Write a program to use the loop to find the factorial of a given number.

The factorial (symbol: $!$) means to multiply all whole numbers from the chosen number down to 1.

For example: calculate the factorial of 5

* `5! = 5 × 4 × 3 × 2 × 1 = 120`

Expected Output:

* `120`

```{r}
N = 5
if(N < 0){
  print("Factorial does not exist for negative numbers")
} else if(N == 0){
  print("The factorial of 0 is 1")
} else{
  for(i in (N-1):1){
    result= N * i
    N = result
  }
}

print(result)
```

## Exercise 14: Reverse a given integer number

Given:

* `76542`

Output:

* `24567`

```{r}
num = 76542
num2 = as.numeric(strsplit(as.character(num), "")[[1]])
list1 = c()
for(i in length(num2):1){
  list1 = c(list1, num2[i])
}
print(as.numeric(paste(list1, collapse = "")))
```

## Exercise 15: Use a loop to display elements from a given list present at odd index positions

Given:

* `my_list = [10, 20, 30, 40, 50, 60, 70, 80, 90, 100]`

Expected output:

* `20 40 60 80 100`

```{r}
my_list = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100)
# Making a vector of the odd indices
list_ind = my_list[seq(0,length(my_list),2)]

# Printing the odd indices
for(i in list_ind){
  print(i)
}
```

## Exercise 16: Calculate the cube of all numbers from 1 to a given number

Write a program to get the cube of all numbers from 1 to a given number

Given:

* `input_number = 6`

```{r}
input_number = 6
for(i in 1:6){
  print(i^3)
}
```

## Exercise 17: Find the sum of the series up to n terms

Write a program to calculate the sum of series up to n term. For example, if n = 5 the series will become 2 + 22 + 222 + 2222 + 22222 = 24690

Given:

* `n = 5`

Expected output:

* `24690`

```{r}
n = 5
# First number of the sequence
start = 2
# Storage for summation
sum_seq = 0

for(i in 1:n){
  # Adding new number to previous number
  sum_seq = sum_seq + start
  start = start * 10 + 2
}

print(sum_seq)
```

## Exercise 18: Print the following pattern

Write a program to print the following start pattern using the for loop:

* `* `
* `* * `
* `* * * `
* `* * * * `
* `* * * * * `
* `* * * * `
* `* * * `
* `* * `
* `*`

```{r}
rows = 5
stor = c()
for(i in 1:(rows-1)){
  for(j in 1:i){
    stor = c(stor, "*")
  }
  print(stor)
  stor = c()
}

for(i in rows:1){
  for(j in 1:i){
    stor = c(stor, "*")
  }
  print(stor)
  stor = c()
}
```

