---
title: "Loop Exercises in R"
author: "Kevin Patyk"
date: "11/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Exercise 1: Write a for loop that iterates over the numbers 1 to 7 and prints the cube of each number using print()**
```{r}
for(i in 1:7){
  print(i^3)
}
```

**Exercise 2: Write a for loop that iterates over the column names of the inbuilt iris dataset and print each together with the number of characters in the column name in parenthesis. Example output: `Sepal.Length (12)`. Use the following functions `print()`, `paste0()` and `nchar()`.**

The difference between `paste()` and `paste0()` is that the argument `sep` by default is ” ” (`paste`) and “” (`paste0`). In conclusion, `paste0()` is faster than `paste()` if our objective is concatenate strings without spaces because we don't have to specify the argument `sep`.
```{r}
for(i in colnames(iris)){
  print(paste0(i, " (", nchar(i), ")"))
}
```

**Exercise 3: Write a while loop that prints out standard random normal numbers (use rnorm()) but stops (breaks) if you get a number bigger than 1.**
```{r}
while(TRUE) {
  x <- rnorm(1)
  print(x)
  if (x > 1) {
    break
  }
}
```

**Exercise 4: Using `next` adapt the loop from last exercise so that doesn’t print negative numbers.**
```{r}
while(TRUE){
  x <- rnorm(1)
  if(x < 0){
    next
  } else if(x > 1){
    break
  } else {
    print(x)
  }
}
```

**Exercise 5: Using a `for` loop simulate the flip a coin twenty times, keeping track of the individual outcomes (1 = heads, 0 = tails) in a vector that you preallocte.**
```{r}
outcomes <- numeric()
for(i in 1:20){
  outcomes[i] <- sample(x = c(1, 0), size = 1)
}
outcomes
```

**Exercise 6: Use a nested `for` loop (a for loop inside a for loop) that produces the following matrix, preallocate the matrix with NA values.**
```{r}
matrix <- matrix(nrow = 5, ncol = 5)
for(i in 1:5){
  for(j in 1:5){
    matrix[i, j] <- abs(i - j)
  }
}
matrix
```

**Exercise 7: Use a `while` loop to investigate the number of terms required before the product 1⋅2⋅3⋅4⋅… reaches above 10 million.**
```{r}
target <- 10^8
product <- 1
n <- 0

while (product <= target) {
  n <- n + 1
  product <- product*n
}
n
```

**Exercise 8: Use a `while` loop to simulate one stock price path starting at 100 and random normally distributed percentage jumps with mean 0 and standard deviation of 0.01 each period. How long does it take to reach above 150 or below 50?**
```{r}
set.seed(22)

price_start <- 100
price <- price_start
target_diff <- 50
n <- 0
while (abs(price_start - price) < target_diff) {
  n <- n + 1
  price <- price*(1 + rnorm(mean = 0, sd = 0.01, n = 1))
}
```

**Exercise 9: Implement a simple version of Guess the number game using a `while` loop. The user should guess a number between 1 and 10, you can use `scan()` to get user input. The loop should break if the user guesses 5.**
```{r eval=FALSE}
x <- 0L
while (TRUE) {
  cat(
    "I am thinking of a number between 1 and 10.",
    "\nTake a guess and press enter twice!"
  )
  x <- scan()[1]
  if (x == 5) {
    cat("Right!")
    break
  }
  cat("Wrong\n")
}
```

**Exercise 10: Implement a multiplication game. A `while` loop that gives the user two random numbers from 2 to 12 and asks the user to multiply them. Only exit the loop after five correct answers. Try using `as.integer(readline())` instead of `scan()` this time.**
```{r eval=FALSE}
total <- 0
while (TRUE) {
  if (!total %in% 0:4) {
    total <- 0L
  }
  n <- sample(c(2:12), 2)
  cat("What is the product of ", n[1], " and ", n[2], "?\n", sep = "")
  x <- as.integer(readline())
  if (x == prod(n)) {
    total <- total + 1L
    cat("Right!")
    if (total == 5) break
    cat(" You just need", 5 - total, "more answers.\n\n")
  } else {
    cat("Wrong\n")
  }
}
```

-----

# End of document

-----

```{r}
sessionInfo()
```

